---
title: 'Covid: Predictions'
author: "Ryan Johnson"
date: "8/25/2021"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(xgboost)
library(tidymodels)
library(modeltime)
library(tidyverse)
library(lubridate)
library(timetk)
library(pins)
```

# Overview

This Rmd forecasts 7-day averages for new cases and new deaths due to Covid-19. Most of the modeling was performed using the `modeltime` package and followed the example outlined [here](https://business-science.github.io/modeltime/articles/getting-started-with-modeltime.html).

The models created in this document **have not been rigorously tested/validated** and are meant primarily to show the basic functionality of the `modeltime` package.

# Get Data and Visualize

```{r data, message=FALSE, warning=FALSE}
# Get data for USA new deaths/caes (7-day averages)
usa_total <- pin_get("ryan/USA_totals_cases_deaths", board = "rsconnect")
  
# Split into Cases and Deaths
## Only model data from past 2 months
usa_total_cases <- usa_total %>% 
  select(Date, total_cases_avg_7) %>% 
  # Fiter by time
  filter(Date >= today() - months(2))

usa_total_deaths <- usa_total %>% 
  select(Date, total_deaths_avg_7) %>% 
  # Fiter by time
  filter(Date >= today() - months(2))


plot_time_series(usa_total_cases, Date, total_cases_avg_7, .interactive = FALSE, .title = "New Cases (7-day avg")
plot_time_series(usa_total_deaths, Date, total_deaths_avg_7, .interactive = FALSE, .title = "New Deaths (7-day avg")
```

# Modeling

## Split Data

```{r}
# Split Data
splits_cases <- initial_time_split(usa_total_cases, prop = 0.9)
splits_deaths <- initial_time_split(usa_total_deaths, prop = 0.9)
```

## Train Models for New Cases

```{r models-cases}
# Model 1: arima_boost ----
model_fit_arima_boosted_c <- arima_boost(min_n = 2,
                                       learn_rate = 0.015) %>%
  set_engine(engine = "auto_arima_xgboost") %>%
  fit(total_cases_avg_7 ~ Date + as.numeric(Date),
      data = training(splits_cases))

# Model 2: ets ----
model_fit_ets_c <- exp_smoothing() %>%
  set_engine(engine = "ets") %>%
  fit(total_cases_avg_7 ~ Date, data = training(splits_cases))

# Model 3: prophet ----
model_fit_prophet_c <- prophet_reg() %>%
  set_engine(engine = "prophet") %>%
  fit(total_cases_avg_7 ~ Date, data = training(splits_cases))

# Model 4: lm ----
model_fit_lm_c <- linear_reg() %>%
  set_engine("lm") %>%
  fit(total_cases_avg_7 ~ as.numeric(Date),
      data = training(splits_cases))

# Model 5: earth ----
model_spec_mars_c <- mars(mode = "regression") %>%
  set_engine("earth")

recipe_spec_c <-
  recipe(total_cases_avg_7 ~ Date, data = training(splits_cases)) %>%
  step_mutate(date_num = as.numeric(Date)) %>%
  step_normalize(date_num) %>%
  step_rm(Date)

wflw_fit_mars_c <- workflow() %>%
  add_recipe(recipe_spec_c) %>%
  add_model(model_spec_mars_c) %>%
  fit(training(splits_cases))
```

## Train Models for New Deaths

```{r models-deaths}
# Model 1: arima_boost ----
model_fit_arima_boosted_d <- arima_boost(min_n = 2,
                                       learn_rate = 0.015) %>%
  set_engine(engine = "auto_arima_xgboost") %>%
  fit(total_deaths_avg_7 ~ Date + as.numeric(Date),
      data = training(splits_deaths))

# Model 2: ets ----
model_fit_ets_d <- exp_smoothing() %>%
  set_engine(engine = "ets") %>%
  fit(total_deaths_avg_7 ~ Date, data = training(splits_deaths))

# Model 3: prophet ----
model_fit_prophet_d <- prophet_reg() %>%
  set_engine(engine = "prophet") %>%
  fit(total_deaths_avg_7 ~ Date, data = training(splits_deaths))

# Model 4: lm ----
model_fit_lm_d <- linear_reg() %>%
  set_engine("lm") %>%
  fit(total_deaths_avg_7 ~ as.numeric(Date),
      data = training(splits_deaths))

# Model 5: earth ----
model_spec_mars_d <- mars(mode = "regression") %>%
  set_engine("earth")

recipe_spec_d <-
  recipe(total_deaths_avg_7 ~ Date, data = training(splits_deaths)) %>%
  step_mutate(date_num = as.numeric(Date)) %>%
  step_normalize(date_num) %>%
  step_rm(Date)

wflw_fit_mars_d <- workflow() %>%
  add_recipe(recipe_spec_d) %>%
  add_model(model_spec_mars_d) %>%
  fit(training(splits_deaths))
```

## Create Model Tables

```{r model-tables}
# Cases
models_tbl_c <- modeltime_table(
    model_fit_arima_boosted_c,
    model_fit_ets_c,
    model_fit_prophet_c,
    model_fit_lm_c,
    wflw_fit_mars_c
)

# Deaths
models_tbl_d <- modeltime_table(
    model_fit_arima_boosted_d,
    model_fit_ets_d,
    model_fit_prophet_d,
    model_fit_lm_d,
    wflw_fit_mars_d
)
```

## Calibrate on testing data

```{r calibrate}
# Calibrate cases
calibration_tbl_c <- models_tbl_c %>%
    modeltime_calibrate(new_data = testing(splits_cases))

# Calibrate deaths
calibration_tbl_d <- models_tbl_d %>%
    modeltime_calibrate(new_data = testing(splits_deaths))
```

## Refit on entire data set

```{r refit}
# Cases
refit_tbl_c <- calibration_tbl_c %>%
    modeltime_refit(data = usa_total_cases)

# Deaths
refit_tbl_d <- calibration_tbl_d %>%
    modeltime_refit(data = usa_total_deaths)
```

## Pin Models
```{r}
# Register Pin Board
pins::board_register_rsconnect(server = Sys.getenv("CONNECT_SERVER"),
                               key = Sys.getenv("CONNECT_API_KEY"))
# Cases
pin(refit_tbl_c,
    name = "Covid_model_cases",
    board = "rsconnect")

# Deaths
pin(refit_tbl_d,
    name = "Covid_model_deaths",
    board = "rsconnect")
```


## Visualize Prediction Models

```{r plot_model, message=FALSE, warning=FALSE}
# Cases: predict 2 weeks
refit_tbl_c %>%
    modeltime_forecast(h = "2 week", actual_data = usa_total_cases) %>%
    plot_modeltime_forecast(
      .legend_max_width = 25, # For mobile screens
      .interactive      = FALSE, 
      .title = "Forecast: Cases"
    )

# Deaths: predict 2 weeks
refit_tbl_d %>%
    modeltime_forecast(h = "2 week", actual_data = usa_total_deaths) %>%
    plot_modeltime_forecast(
      .legend_max_width = 25, # For mobile screens
      .interactive      = FALSE,
      .title = "Forecast: Deaths"
    )
```
